{\fontfamily{ptm}\selectfont Linear Convergence of Accelerated Gradient without Restart}

### 2025-08-27 Notebook LaTEX/Formalized Notes/Notes on QSCNVX/main.tex
;~;, we cry

\section{Precursors materials for our proofs of linear convergence}
    The following two definitions defines the accelerated proximal gradient algorithm. 
    % DEFINITION =======================================================================================================
    \begin{definition}[similar triangle form of accelerated proximal gradient]\;\label{def:st-apg}\\
        The definition is about $((\alpha_k)_{k \ge 0}, (q_k)_{k \ge 0}, (B_k)_{k \ge 0}, (y_k)_{k \ge 0}, (x_k)_{k \ge -1}, (v_k)_{k \ge -1})$. 
        These sequences satisfy:
        \begin{enumerate}[nosep]
            \item $x_{-1}, y_{- 1}\in \RR^n$ are arbitrary initial condition of the algorithm;
            \item $(q_k)_{k \ge 1}$ be a sequence such that $q_k \in [0, 1)$ for all $k \ge 1$;
            \item $(\alpha_k)_{k \ge 1}$ be a sequence such that $\alpha_0 \in (0, 1]$, and for all $k \ge 1$ it has $\alpha_k \in (q_k, 1)$;
            \item $(B_k)_{k \ge 0}$ has $B_k \ge 0$, it's a nonnegative sequence. 
        \end{enumerate}
        Then an algorithm satisfies the similar triangle form of Nesterov's accelerated gradient if it generates iterates $(y_k, x_k, v_k)_{k \ge 1}$ such that for all $k\ge 0$: 
        \begin{align*}
            y_k &= \left(\frac{\alpha_k - q_k}{1 - q_k}\right)v_{k - 1} 
            + \left(\frac{1 - \alpha_k}{1 - q_k}\right) x_{k - 1},
            \\
            x_k &= T_{B_k}(y_k), D_f(x_k, y_k) \le \frac{B_k}{2}\Vert x_k - y_k\Vert^2, 
            \\
            v_k &= x_{k - 1} + \alpha_k^{-1}(x_k - x_{k - 1}). 
        \end{align*}
    \end{definition}
    % DEFINITION =======================================================================================================
    \begin{definition}[relaxed momentum sequence]\label{def:rlx-momentum-seq}
        The following definition is about sequences $((\alpha_k)_{k \ge 0}, (q_k)_{k \ge 0}, (\rho_k)_{k \ge 0})$. 
        Let 
        \begin{enumerate}[nosep]
            \item $(q_k)_{k \ge 0}$ is a sequence such that $q_k \in [0, 1)$ for all $k \ge 0$;
            \item $(\alpha_k)_{k \ge 0}$ be such that $\alpha_0 \in (0, 1]$, and for all $k \ge 1$ it has $\alpha_k \in (q_k, 1)$;
            \item $(\rho_k)_{k \ge 0}$ is a strictly positive sequence for all $k \ge 1$. 
        \end{enumerate}
        The sequences $q_k, \alpha_k$ are considered relaxed momentum sequence if for all $k \ge 1$ it satisfies the relation that: 
        \begin{align*}
            \rho_{k - 1} &= \frac{\alpha_k(\alpha_k - q_k)}{(1 - \alpha_{k})\alpha_{k - 1}^2}. 
        \end{align*}
    \end{definition}
    % DEFINITION =======================================================================================================
    \begin{definition}[proximal gradient gap]\label{def:pg-gap}
        Let $F = f + g$ where $f$ is $L$ Lipschitz smooth and $g$ is convex. 
        Then the proximal gradient mapping $T_B(x) = \hprox_{B^{-1}g}(x -  B^{-1}\nabla f(x))$ is a singleton, and $\dom T_B = \RR^n$. 
        Let $\mu, B$ be parameters such that $B > \mu \ge 0$. 
        We define the proximal gradient gap $\mathcal E(z, y, \mu, B)$ is the mapping: 
        \begin{align*}
            \mathcal E(z, y, \mu, B) &:= 
            F(z) - F(T_B(y)) 
            - \left\langle B(y - T_B(y)), z - y\right\rangle
            - \frac{\mu}{2}\Vert z - y\Vert^2
            - \frac{B}{2}\Vert y - T_B(y)\Vert^2. 
        \end{align*}
    \end{definition}
    \begin{remark}
        This expression is the same as the proximal gradient inequality up to a negative sign, after moving everything to one side. 
    \end{remark}

\section{Deriving the convergence rate}
    To derive the convergence rate of algorithm satisfying Definition \ref{def:st-apg}, \ref{def:rlx-momentum-seq}, we leverage Definition \ref{def:pg-gap}. 
    The first two subsections prepare for the results and the the convergence results are derived by the end. 
    \par
    The following assumption is about the Proximal Gradient gap, it's required to obtain our first result. 
    % ASSUMPTION =======================================================================================================
    \begin{assumption}[generic assumptions for convergence]\label{ass:for-cnvg}
        The following assumption is about $(F, f, g, \mathcal E, \mu, L)$, it is the configuration needed to derive the convergence rate of algorithms that satisfy Definition \ref{def:st-apg}. 
        We assume that there exists $\mu \ge 0$ such that the followings are true: 
        \begin{enumerate}[nosep]
            \item Let $F = f + g$ where $f$ is $L$ Lipschitz smooth and, $g$ is closed convex and proper.
            \item $\forall y \in \RR^n, \exists B \ge 0\; \exists \bar y$ such that $\mathcal E(\bar y, y, \mu, B) \ge 0$. 
            \item For all $z, y \in \RR^n$, there exists $B > \mu$ such that $\mathcal E(z, y, \mu, B) + \frac{\mu}{2}\Vert z - y\Vert^2 \ge 0$.   
        \end{enumerate}
    \end{assumption}
    \begin{remark}
        Note that:
        \begin{enumerate}
            \item If the function is convex, all conditions are satisfies for $\mu = 0$, and for all $\bar y \in \RR^n$. 
            \item If $\mu \ge 0$ satisfies item (ii), (iii), then it's also satisfies for all $\tilde \mu$ such that $0 \le \tilde \mu \le \mu$. For the best convergence rate, the largest such $\mu$ is of our interest. 
        \end{enumerate}

@@

        Using the the above and the update rule for $v_k$ in Definition \ref{def:st-apg}. 

@@

            (1 - q_k)^{-1}y_k &= 

@@

            (\alpha_k - q_k)x_{k - 2} 
            + \alpha_{k - 1}^{-1}(x_{k - 1} - x_{k - 2}) + (1 - \alpha_k)x_{k - 1}

@@

        Multiply $(1 - q_k)$ on both sides yield: 

@@

                    (\alpha_{k - 1}^{-1} -1)(\alpha_k - q_k)g 

@@

                \item The sequences $(\alpha_k)_{k\ge 0}, (\rho_{k})_{k \ge 0}, (q_k)_{k \ge 0}$ satisfy Definition \ref{def:rlx-momentum-seq}. 

@@

                -\frac{B_k(1 - \alpha_k)\rho_{k - 1}\alpha_{k - 1}^2}{2}\Vert \bar x - v_{k - 1}\Vert^2

@@

            Subtracting $-\frac{B_k(1 - \alpha_k)\rho_{k - 1}\alpha_{k - 1}^2}{2}\Vert \bar x - v_{k - 1}\Vert^2$ from both sides, the coefficient for $\Vert \bar x - v_{k - 1}\Vert^2$ comes out to be: 

@@

                - \frac{B_k(1 - \alpha_k)\rho_{k - 1}\alpha_{k - 1}^2}{2} 

@@

                    \alpha_k^2 + (1 - \alpha_k)\rho_{k - 1}\alpha_{k - 1}^2

@@

            Suppose that all of the following are satisfied

@@

                \item The sequences $(\alpha_k)_{k \ge 0}, (\rho_k)_{k \ge 0}, (q_k)_{k \ge 0}$ satisfies Definition \ref{def:rlx-momentum-seq}. 

@@

                + \frac{B_k(1 - \alpha_k)\rho_{k - 1}\alpha_{k - 1}^2}{2}\Vert \bar x_k - v_{k - 1}\Vert^2. 

@@

                - \frac{B_k(1 - \alpha_k)\rho_{k - 1}\alpha_{k - 1}^2}{2}\Vert \bar x_k - v_{k - 1}\Vert^2

@@

            At (1), we used Lemma \ref{lemma:cnvg-prep-part3}, and substituted $\Xi_k$. 

@@

            At (2), we used the $\mathcal E(\bar x_k, y_k, \mu, B_k)\le 0$ by chosing $\bar x_k = \bar y$ in Assumption \ref{ass:for-cnvg} (iii) to make the inequality. 

@@

            At this point, we had proved what we wanted becaues using the definitions of $\Xi_k, \Xi_k'$ it has:

@@

        Our convergence rate is expressed for sequences $(\alpha_k)_{k \ge 0}, (\rho_k)_{k\ge0}$ as long as they satisfies Definition \ref{def:rlx-momentum-seq}. 

@@

        The following assumptions describes the behaviors of an algorithm satisfying Definition \ref{def:st-apg}, its parameters, and the properties of the objective function. 

@@

            Denote $\beta_k = \prod_{i = 1}^k(1 - \alpha_i) \max \left(\frac{B_i\rho_{i - 1}}{B_{i - 1}}, 1\right)$, $\beta_0 = 1$. 

@@

                \right)\left(
                    F(x_{k- 1}) - F(\bar x) + \frac{B_k\alpha_{k -1}^2 \rho_{k - 1}}{2}\Vert \bar x - v_{k - 1}\Vert^2

@@

@@

                \right)\left(
                    F(x_{k- 1}) - F(\bar x) + 
                    \frac{B_k\rho_{k - 1}}{B_{k - 1}}\frac{B_{k - 1}\alpha_{k -1}^2}{2}\Vert \bar x - v_{k - 1}\Vert^2

@@

                \right)\max\left(

@@

                    F(x_{k- 1}) - F(\bar x) + 
                    \frac{B_{k - 1}\alpha_{k -1}^2 }{2}\Vert \bar x - v_{k - 1}\Vert^2

@@

            From the above, an recurrence relation is formed for $k \ge 1$, unrolloing the recurrence relation it has 

@@

                \right)\max\left(

@@

                    F(x_{k- 1}) - F(\bar x) + 
                    \frac{B_{k - 1}\alpha_{k -1}^2 }{2}\Vert \bar x - v_{k - 1}\Vert^2

@@

        The above theorem shows that the convergence rate is exclusively depeneded on the momentum sequence. 

@@

@@

    In this sections, we exam Assumption \ref{ass:lin-cnvg} and, propose examples for it. 
    In Necoara et al.'s setting, they have the following Assumptoins for their objective function. 

@@

            \item $f: X\rightarrow \RR$ has $L$ Lipschitz continous gradient, and it's convex. 

@@

    By... the above definition satisfies Assumption \ref{ass:for-cnvg} 

@@

        Therefore, it satisfies Assumtpion \ref{ass:for-cnvg}. 

@@

        Denote $x^+ = T_B(x)$, since $x^+$ is a minimizer therefore it has $\forall x \in X\; h(z) - h(x^+) \ge \frac{B}{2}\Vert z - x^+\Vert^2$. 

@@

        The above results shows that Assumption \ref{ass:for-cnvg} has been satisfied, with $\bar y = \Pi_{X^+}y$. 

@@

    \subsection{SCNVX affine composite over simple cone}

>>>>

\section{Precursors materials for our proofs of linear convergence}
    The following two definitions defines the accelerated proximal gradient algorithm. 
    % DEFINITION =======================================================================================================
    \begin{definition}[similar triangle form of accelerated proximal gradient]\;\label{def:st-apg}\\
        The definition is about $((\alpha_k)_{k \ge 0}, (q_k)_{k \ge 0}, (B_k)_{k \ge 0}, (y_k)_{k \ge 0}, (x_k)_{k \ge -1}, (v_k)_{k \ge -1})$. 
        These sequences satisfy:
        \begin{enumerate}[nosep]
            \item $x_{-1}, y_{- 1}\in \RR^n$ are arbitrary initial condition of the algorithm;
            \item $(q_k)_{k \ge 1}$ be a sequence such that $q_k \in [0, 1)$ for all $k \ge 1$;
            \item $(\alpha_k)_{k \ge 1}$ be a sequence such that $\alpha_0 \in (0, 1]$, and for all $k \ge 1$ it has $\alpha_k \in (q_k, 1)$;
            \item $(B_k)_{k \ge 0}$ has $B_k \ge 0$, it's a nonnegative sequence. 
        \end{enumerate}
        Then an algorithm satisfies the similar triangle form of Nesterov's accelerated gradient if it generates iterates $(y_k, x_k, v_k)_{k \ge 1}$ such that for all $k\ge 0$: 
        \begin{align*}
            y_k &= \left(\frac{\alpha_k - q_k}{1 - q_k}\right)v_{k - 1} 
            + \left(\frac{1 - \alpha_k}{1 - q_k}\right) x_{k - 1},
            \\
            x_k &= T_{B_k}(y_k), D_f(x_k, y_k) \le \frac{B_k}{2}\Vert x_k - y_k\Vert^2, 
            \\
            v_k &= x_{k - 1} + \alpha_k^{-1}(x_k - x_{k - 1}). 
        \end{align*}
    \end{definition}
    % DEFINITION =======================================================================================================
    \begin{definition}[relaxed momentum sequence]\label{def:rlx-momentum-seq}
        The following definition is about sequences $((\alpha_k)_{k \ge 0}, (q_k)_{k \ge 0}$. 
        They satisfy:
        \begin{enumerate}[nosep]
            \item $(q_k)_{k \ge 0}$ is a sequence such that $q_k \in [0, 1)$ for all $k \ge 0$;
            \item $(\alpha_k)_{k \ge 0}$ has $\alpha_k \in (0, 1]$, and for all $k \ge 0$. 
        \end{enumerate}
    \end{definition}
    % DEFINITION =======================================================================================================
    \begin{definition}[proximal gradient gap]\label{def:pg-gap}
        Let $F = f + g$ where $f$ is $L$ Lipschitz smooth and $g$ is convex. 
        Then the proximal gradient mapping $T_B(x) = \hprox_{B^{-1}g}(x -  B^{-1}\nabla f(x))$ is a singleton, and $\dom T_B = \RR^n$. 
        Let $\mu, B$ be parameters such that $B > \mu \ge 0$. 
        We define the proximal gradient gap $\mathcal E(z, y, \mu, B)$ is the mapping: 
        \begin{align*}
            \mathcal E(z, y, \mu, B) &:= 
            F(z) - F(T_B(y)) 
            - \left\langle B(y - T_B(y)), z - y\right\rangle
            - \frac{\mu}{2}\Vert z - y\Vert^2
            - \frac{B}{2}\Vert y - T_B(y)\Vert^2. 
        \end{align*}
    \end{definition}
    \begin{remark}
        This expression is the same as the proximal gradient inequality up to a negative sign, after moving everything to one side. 
    \end{remark}

\section{Deriving the convergence rate}
    To derive the convergence rate of algorithm satisfying Definition \ref{def:st-apg}, \ref{def:rlx-momentum-seq}, we leverage Definition \ref{def:pg-gap}. 
    The first two subsections prepare for the results and the convergence results are derived by the end. 
    \par
    The following assumption is about the Proximal Gradient gap, it's required to obtain our first result. 
    % ASSUMPTION =======================================================================================================
    \begin{assumption}[generic assumptions for convergence]\label{ass:for-cnvg}
        The following assumption is about $(F, f, g, \mathcal E, \mu, L)$, it is the configuration needed to derive the convergence rate of algorithms that satisfy Definition \ref{def:st-apg}. 
        We assume that there exists $\mu \ge 0$ such that the followings are true: 
        \begin{enumerate}[nosep]
            \item Let $F = f + g$ where $f$ is $L$ Lipschitz smooth and, $g$ is closed convex and proper.
            \item $\forall y \in \RR^n, \exists B \ge 0\; \exists \bar y$ such that $\mathcal E(\bar y, y, \mu, B) \ge 0$. 
            \item For all $z, y \in \RR^n$, there exists $B > \mu$ such that $\mathcal E(z, y, \mu, B) + \frac{\mu}{2}\Vert z - y\Vert^2 \ge 0$.   
        \end{enumerate}
    \end{assumption}
    \begin{remark}
        Note that:
        \begin{enumerate}
            \item If the function $f$ is convex and smooth, $g$ is convex then, all conditions are satisfied for $\mu = 0$, and for all $\bar y \in \RR^n$. 
            \item If $\mu \ge 0$ satisfies item (ii), (iii), then it also satisfies for all $\tilde \mu$ such that $0 \le \tilde \mu \le \mu$. For the best convergence rate, the largest such $\mu$ is of our interest. 
        \end{enumerate}

@@

        Using the above and the update rule for $v_k$ in Definition \ref{def:st-apg}. 

@@

            (1 - q_k)y_k &= 

@@

            (\alpha_k - q_k)((1 - \alpha_{k - 1}^{-1})x_{k -2} + \alpha_{k - 1}^{-1}x_{k - 1}) 

@@

        Divide by $(1 - q_k)$ on both sides yield: 

@@

@@

                \item The sequences $(\alpha_k)_{k\ge 0}, (q_k)_{k \ge 0}$ satisfy Definition \ref{def:rlx-momentum-seq}. 

@@

                - \frac{B_k\alpha_k(\alpha_k - q_k)}{2}\Vert \bar x - v_{k - 1}\Vert^2

@@

            Subtracting $\frac{B_k\alpha_k(\alpha_k - q_k)}{2}\Vert \bar x - v_{k - 1}\Vert^2$ from both sides, the coefficient of $\Vert \bar x - v_{k - 1}\Vert^2$ comes out to be: 

@@

                - \frac{B_k\alpha_k(\alpha_k - q_k)}{2}

@@

                    \alpha_k^2 - \alpha_k(\alpha_k - q_k)

@@

            Suppose that all the following are satisfied

@@

                \item The sequences $(\alpha_k)_{k \ge 0}, (q_k)_{k \ge 0}$ satisfies Definition \ref{def:rlx-momentum-seq}. 

@@

                + \frac{B_k\alpha_k(\alpha_k - q_k)}{2}\Vert \bar x_k - v_{k - 1}\Vert^2. 

@@

                - \frac{B_k\alpha_k(\alpha_k - q_k)}{2}\Vert \bar x_k - v_{k - 1}\Vert^2

@@

            At (1), we used Lemma \ref{lemma:cnvg-prep-part1}, and substituted $\Xi_k$. 

@@

            At (2), we used the $\mathcal E(\bar x_k, y_k, \mu, B_k)\le 0$ by choosing $\bar x_k = \bar y$ in Assumption \ref{ass:for-cnvg} (iii) to make the inequality. 

@@

            At this point, we had proved what we wanted because using the definitions of $\Xi_k, \Xi_k'$ it has:

@@

        Our convergence rate is expressed for sequences $(\alpha_k)_{k \ge 0}, (q_k)_{k \ ge 0}$ if they satisfy Definition \ref{def:rlx-momentum-seq}. 

@@

        The following assumptions describe the behaviors of an algorithm satisfying Definition \ref{def:st-apg}, its parameters, and the properties of the objective function. 

@@

            Denote 
            $$
                \beta_k = \prod_{i = 1}^k\max\left(
                    1 - \alpha_i, 
                    \frac{B_i\alpha_i(\alpha_i - q_i)}{\alpha_{i - 1}^2B_{i - 1}}
            $$

@@

                \right)(F(x_{k- 1}) - F(\bar x))
                + \frac{(1 - \alpha_k)B_k\alpha_{k -1}^2 \rho_{k - 1}}{2}\Vert \bar x - v_{k - 1}\Vert^2

@@

                &\underset{\text{(1)}}{=} 

@@

                \right)(F(x_{k- 1}) - F(\bar x))
                + \frac{\alpha_k(\alpha_k - q_k)B_k\alpha_{k - 1}^2}{\alpha_{k - 1}^2}\Vert \bar x - v_{k - 1}\Vert^2

@@

                    1 - \alpha_k, 
                    \frac{B_i\alpha_k(\alpha_k - q_k)}{\alpha_{k - 1}^2B_{k - 1}}

@@

                    F(x_{k - 1}) - F(\bar x)
                    + \frac{B_{k - 1}\alpha_{k - 1}^2}{2}\Vert \bar x - v_{k - 1}\Vert^2

@@

            From the above, a recurrence relation is formed for $k \ge 1$, unrolling the recurrence relation it has 

@@

                \\&\le 
                \prod_{i = 1}^k\max\left(
                    1 - \alpha_i, 
                    \frac{B_i\alpha_i(\alpha_i - q_i)}{\alpha_{i - 1}^2B_{i - 1}}

@@

                    F(x_0) - F(\bar x)
                    + \frac{B_0\alpha_0^2}{2}\Vert \bar x - v_0\Vert^2

@@

        \textcolor{red}{
            SOMETHING IS WRONG, LET US EXPLAIN.
            The parameter $q_k = \mu/B_k$ cannot be eliminated from Definition \ref{def:st-apg}. 
            If one were to implement any algorithm without the exact knowledge of $\mu$, then the convergence results we just derived are not applicable. 
            Implementations of the algorithm, and the convergence results requires the presence of both parameters $q_k, \alpha_k$. 
            Doomed. 
            Because in the applications of our interests, the exact knowledge of $\mu$ is impossible to know. 

@@

        \end{proof}
    \subsection{Parameter free representations}

@@

    We will check Assumption \ref{ass:lin-cnvg} and, propose examples for it.
    In Necoara et al.'s setting, they have the following assumptions for their objective function. 

@@

            \item $f: X\rightarrow \RR$ has $L$ Lipschitz continuous gradient, and it's convex. 

@@

@@

        Therefore, it satisfies Assumption \ref{ass:for-cnvg}. 

@@

        Denote $x^+ = T_B(x)$, since $x^+$ is a minimizer therefore it has $(\forall x \in X)\; h(z) - h(x^+) \ge \frac{B}{2}\Vert z - x^+\Vert^2$. 

@@

        The above results showed that Assumption \ref{ass:for-cnvg} (ii) has been satisfied, with $\bar y = \Pi_{X^+}y$. 
        To see (iii) in Assumption \ref{ass:for-cnvg}, observe that Assumption \ref{ass:necoara} has $f$ is convex. 

@@

    \subsection{A nonconvex example}

    \subsection{Examples of Q-SCNVX}
